Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-04-13 manuscript/manuscript.tex
Changed every todo note to inline todo note

\citet{Uieda2016} developed a forward gravitational model based on 
tesseroids with homogeneous densities using the GLQ approximation 
method.
The main issue of the GLQ is that the approximation becomes less 
accurate when the computation point is closer to the tesseroid of for 
smaler GLQ order \citep{Ku1977}.
Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
ensures the accuracy implementing a modified version of the adaptive 
discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain 
distance-size ratio is lower than a predefined value ($D$) that controls the 
accuracy of the computation.
\citet{Uieda2016} have also objectively obtained standard values of $D$ 
for gravitational potential, gradients and tensor components comparing 
the numerical model with the fields generated by a spherical shell, 
which constitutes a special case of the volume integrals that has 
analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013} 
\todo{citar mas??}.

@@

Furthermore, we have the intention to include it into a future release 
of the Fatiando a Terra library in order to make its distribution and 
maintenance easier.
\todo{include this?}

@@

From equation \ref{eq:glq-var-dens} we can observe that the GLQ 
approximates the gravitational fields of a tesseroid as the ones 
generated by $N_\lambda N_\phi N_r$ point masses located at the nodes 
of the Legendre polynomials, where $N_i$ are the orders of the 
quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$, 
in the same way as the GLQ works in the homogeneous density case.
Nevertheless, in this new variable density scenario, the information of 
the density function is summarised into the weights by taking into 
account only the values it assumes on those nodes.
Although this may sound discouraging, we intent to prove otherwise and 
show that this method, along with well suited discretization 
algorithms, approximates the gravitational fields with good accuracy.
\todo{no se si poner esta ultima oracion}

@@

We will perform similar tests, but now with variable density spherical 
shell.
From the infinite density functions that can be proposed, we have 
chosen three characteristic ones: a linear, an exponential and a 
discontinuous function.
\todo{sacaria todo este parrafo y lo introduciria mas adelante}

@@

For future purposes we are going to obtain the gravity potential for 
three different functions from the perspective of their variation in 
$r'$: a linear and an exponential density function.
\todo{lo sacaria}

@@

We can see the results for the potential, the vertical component of the 
gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on 
Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
We exclude the results corresponding to the other two diagonal 
components of the tensor because they are similar to the ones obtained 
for the $g_{zz}$, although they can be found on the repository. 
\todo{chequear}

@@

It worth mention that a greater $D$ or a lower $\delta$ 
increases the number of discretizations of the tesseroid.
Although a greater $D$ may divide it in the latitude, longitude 
and/or radial direction; a lower $\delta$ will only discretize the 
original tesseroid in the radial direction.
This means that a decrease in $\delta$ will produce a more efficient 
computation than an increase in $D$.
\todo{Should we keep this paragraph? The same is explained bellow.}

@@

We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth.
It is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}

@@

On Figures \ref{fig:speed-comparison}a and \ref{fig:speed-comparison}b we can see the results of these comparison.
In the linear density case the density-based discretization algorithm is not applied because the inequation \ref{eq:delta-density} is never satisfied.
For this reason, the homogeneous density model is up to 5 times faster than the variable density one in this case.
On the other hand, the density-based discretization algorithm makes the computation slower: up to 20 times slower than the homogeneous density code.
Nevertheless, this increase in the computation time pays off with a guarantee\todo{guarantee is a noun?} on its accuracy.

@@

We must also define a density function for the entire model.
\citet{Sigismondi2012} proposed a minimum and maximum density contrast for the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively. \todo{ver que esos valores son de contraste y tienen q ir negativos. poner los valores absolutos.}
So we have chosen, for this specific example, an exponential density (like the one in equation \ref{eq:density-exp}) that assumes the minimum value on the surface and the maximum at 5858m, i.e. the bottom of the basin, with a value of $b$ equal to 3km.

>>>>

\citet{Uieda2016} developed a forward gravitational model based on 
tesseroids with homogeneous densities using the GLQ approximation 
method.
The main issue of the GLQ is that the approximation becomes less 
accurate when the computation point is closer to the tesseroid of for 
smaler GLQ order \citep{Ku1977}.
Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
ensures the accuracy implementing a modified version of the adaptive 
discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain 
distance-size ratio is lower than a predefined value ($D$) that controls the 
accuracy of the computation.
\citet{Uieda2016} have also objectively obtained standard values of $D$ 
for gravitational potential, gradients and tensor components comparing 
the numerical model with the fields generated by a spherical shell, 
which constitutes a special case of the volume integrals that has 
analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013} 
\todo[inline]{citar mas??}.

@@

Furthermore, we have the intention to include it into a future release 
of the Fatiando a Terra library in order to make its distribution and 
maintenance easier.
\todo[inline]{include this?}

@@

From equation \ref{eq:glq-var-dens} we can observe that the GLQ 
approximates the gravitational fields of a tesseroid as the ones 
generated by $N_\lambda N_\phi N_r$ point masses located at the nodes 
of the Legendre polynomials, where $N_i$ are the orders of the 
quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$, 
in the same way as the GLQ works in the homogeneous density case.
Nevertheless, in this new variable density scenario, the information of 
the density function is summarised into the weights by taking into 
account only the values it assumes on those nodes.
Although this may sound discouraging, we intent to prove otherwise and 
show that this method, along with well suited discretization 
algorithms, approximates the gravitational fields with good accuracy.
\todo[inline]{no se si poner esta ultima oracion}

@@

We will perform similar tests, but now with variable density spherical 
shell.
From the infinite density functions that can be proposed, we have 
chosen three characteristic ones: a linear, an exponential and a 
discontinuous function.
\todo[inline]{sacaria todo este parrafo y lo introduciria mas adelante}

@@

For future purposes we are going to obtain the gravity potential for 
three different functions from the perspective of their variation in 
$r'$: a linear and an exponential density function.
\todo[inline]{lo sacaria}

@@

We can see the results for the potential, the vertical component of the 
gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on 
Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
We exclude the results corresponding to the other two diagonal 
components of the tensor because they are similar to the ones obtained 
for the $g_{zz}$, although they can be found on the repository. 
\todo[inline]{chequear}

@@

It worth mention that a greater $D$ or a lower $\delta$ 
increases the number of discretizations of the tesseroid.
Although a greater $D$ may divide it in the latitude, longitude 
and/or radial direction; a lower $\delta$ will only discretize the 
original tesseroid in the radial direction.
This means that a decrease in $\delta$ will produce a more efficient 
computation than an increase in $D$.
\todo[inline]{Should we keep this paragraph? The same is explained bellow.}

@@

We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth.
It is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo[inline]{agregar repo, vale la pena poner el link?}

@@

On Figures \ref{fig:speed-comparison}a and \ref{fig:speed-comparison}b we can see the results of these comparison.
In the linear density case the density-based discretization algorithm is not applied because the inequation \ref{eq:delta-density} is never satisfied.
For this reason, the homogeneous density model is up to 5 times faster than the variable density one in this case.
On the other hand, the density-based discretization algorithm makes the computation slower: up to 20 times slower than the homogeneous density code.
Nevertheless, this increase in the computation time pays off with a guarantee\todo[inline]{guarantee is a noun?} on its accuracy.

@@

We must also define a density function for the entire model.
\citet{Sigismondi2012} proposed a minimum and maximum density contrast for the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively. \todo[inline]{ver que esos valores son de contraste y tienen q ir negativos. poner los valores absolutos.}
So we have chosen, for this specific example, an exponential density (like the one in equation \ref{eq:density-exp}) that assumes the minimum value on the surface and the maximum at 5858m, i.e. the bottom of the basin, with a value of $b$ equal to 3km.
